---
title: "Homework5"
author: "aa2569@scarletmail.rutgers.edu"
date: "10/10/2022"
output: html_document
---
  
## Decompose
  
* Data Set used - cafe

```{r}
library(fpp)
library(fpp2)

head(cafe)
plot(cafe)
stl_decomp <- stl(cafe,s.window ="periodic")
plot(stl_decomp)
attributes(stl_decomp)
```

* Decompose function is used to separate the the trend, seasonality and the irregularity in the data.
* From the plot above, we can see that there is a seasonality in the data and there is an upward trend.
* Towards the end of the data, the irregulairty component is observed to be increasing.

```{r}
seasadj(stl_decomp)

plot(cafe)
lines(seasadj(stl_decomp), col="Red")
```

* The seasadj function adjusts the seasonality of the data and gives the data without the seasonal component.
* The plot above shows the data without seasonal component in Red. 

```{r}
f_stl <- forecast(stl_decomp,h=15)
f_stl
plot(f_stl)
```

* The f_stl above is forecasting the data. the "h" mentioned is the time period for which we need the forecast. 
* In the plot, the dark and light gray area is the 85% and 95% confidence interval of the forecast respectively.

```{r}
f_stl2 <- forecast(stl_decomp,h=100)
plot(f_stl2)
```

* Here, the h is taken to be 100 to understand fucnction better.
* This shows us that as time period of forecast increases, the confidence interval also increases.
* Which means as the more time we forecast, the more worse our forecast is. Which makes sense.

## Another way of Decompose

```{r}
decomp_cafe <- decompose(cafe)

attributes(decomp_cafe)
seasadj(decomp_cafe)
```

* decompose() function is one more way of using decompose.
* In the attribute we can see that there are trend, seasonal, and random variations.
* seasadj does the same job of adjusting the seasonal variation in the data.
* Unlike the previous decompose function, this decompose is slightly different which is shown below.

```{r}
plot(decomp_cafe)
```

* In the plot, it can be seen that the order of showing the trend, seasonal, and irregularity is different but the visuals are the same.
* "random" indicates the irregularity component and it is shown is a line graph for this function.
