---
title: "Homework 3"
author: "aa2569@scarletmail.rutgers.edu"
date: "9/26/2022"
output: html_document
---
## Time Series - Cafe (fpp)

* The time series from fpp package shows the time series values of quarterly expenditure on cafes, restaurants and takeaway food services in Australia.
* The data shows the values from Qtr2 of 1982 to Qtr4 of 2010.
* The units of the values given have not been mentioned. By looking at the data, assuming the data to be in million dollars.


## Plotting the Time Series

```{r}
library(fpp)
plot(cafe,main = "Quarterly expenses for eating outside in Asutralia", xlab = "Time", ylab="Expenses for eating outside")
```

* The plot shows a trend going upwards with time.
* It appears to have some seasonality as well. We can see a peak at 3rd quarter every year and a dip for the preceding first quarter.
* As per the observations, the ACF plot should have the following features:
    + The ACF should have all significant values.
    + The correlation values gradually come down with lag.
    + There might be a wave visual of the correlation depicting the seasonal change in the data.
    

## ACF Plot

```{r}
Acf(cafe, main = "ACF plot - Cafe")
```


* The ACF plot shows all the correlation values above the 95% confidence interval.
* This shows that the historic data has significant impact on the preceding data.
* The predicted seasonality is not clearly visible in the ACF plot.
* A very slight wave nature can be seen, but is not so clear. So, our prediction of the data showing seasonality cannot be justified.


## PACF Plot

```{r}
Pacf(cafe, main = "PACF plot - Cafe")
```

* The PACF plot shows two lines crossing the significance line.
* By the above two inputs of ACF and PACF, we can conclude that we need to use AR(2) model for forecasting the trend.


## Holt-Winters Filtering

```{r}
cafe_hw1 <- HoltWinters(cafe)
plot(cafe_hw1)
```

* It can be observed that the Holt winters fit is deviating at multiple points.
* The deviations are also massive compared to the original data.
* This indicates that the Holt winters filtering is wrong in this case, and the alpha, beta, gamma values need to be updated compared to the default ones.

```{r}
cafe_hw1f <- forecast(cafe_hw1)
plot(cafe_hw1f$residuals, main = "Residuals plot")
```

* The residuals seem to have an increasing trend as the time passes.
* This shouldn't be the case as the residuals need to be unpredictable for the forecast to be a good fit.

```{r}
hist(cafe_hw1f$residuals, main = "Histogram Plot")
```

* The plot ideally should be a perfect normal distribution contentrated at zero.
* In this case, the plot is not a perfect normal distribution stating the forecast to be wrong.

```{r}
Acf(cafe_hw1f$residuals, main = "ACF Residuals")
```

* Around 3 lines almost reached the confidence internal line.
* Two lines have crossed the confidence internal - meaning there is significant difference and the residuals can be predictable.

```{r}
accuracy(cafe_hw1f)
```

* The ME and RMSE are extremely high as expected.
* For the forecast to be the best, the RMSE should be the least.
* As the results from above histogram, ACF suggested, the accuracy test justified that the Holt-Winters forecast is not the right fit in this case.


## Exponential Smoothing Curves - Testing Additive, Multiplicative, and Damped

```{r}
#Additive
cafe_hw101 <- hw(cafe, seasonal = "additive")
summary(cafe_hw101)
plot(cafe_hw101)
```


```{r}
#Multiplicative
cafe_hw102 <- hw(cafe, seasonal = "multiplicative")
summary(cafe_hw102)
plot(cafe_hw102)
```


```{r}
#Damped+Multiplicative
cafe_hw103 <- hw(cafe, damped = TRUE, seasonal = "multiplicative")
summary(cafe_hw103)
plot(cafe_hw103)
```

* The hw functions for additive, multiplicative, and damped multiplicative are tested.
* Out of all the three, Multiplicative graph has a better RMSE. But, even that is above 122 suggesting it not be a good fit.


**Summary**

* As discussed in the class, Holt-Winters works well only if the data is a combination of Trend and Seasonal time series.
* In this case, the data showed only the characteristics of a Trend time series and not a Seasonal time series.
* Out of the hw functions, Multiplicative forecast showed the best results compared to the other two.