---
title: "BF_HW_ARIMA"
author: "aa2569@scarletmail.rutgers.edu"
date: "11/10/2022"
output: html_document
---
  
```{r}
library(fpp)
library(fpp2)
library(forecast)
```

## Data set - cafe (fpp)

#### Cafe plot

```{r}
plot(cafe, main = 'Quarterly expenditure on eating out in Australia', xlab = 'Year', ylab = 'Expenditure')
```

* We can see that the data has both trend and seasonality.
* Upon testing various forecast methods earlier, we have seen an increase in the residuals towards the end everytime.
* Which signifies that the latest data is quite different compared to the data in the past.
* So, here we going to consider a window of the data from 2005.

#### Window Plot

```{r}
cafe_win <- window(cafe, start = 2005)
plot(cafe_win, main = 'Quarterly expenditure on eating out in Australia', xlab = 'Year', ylab = 'Expenditure')
```

* Even here, we can observe both trend and seasonality from the first glance.
* We can check this with an Acf function.

#### Acf plot

```{r}
Acf(cafe_win, main = 'Acf Plot - Cafe from 2005')
```

* The Acf plot shows both trend and seasonality presense here.
* As there is seasonality we use both nsdiffs and ndiffs.



## ARIMA for data with seasonal component

#### Checking the Differencing needed for seasonal component

```{r}
nsdiffs(cafe_win) #nsdiffs is for seasonal
```

* nsdiffs is use to know how many differences we need to do for seasonality component.
* We first go ahead with seasonality beacuse most of the cases, ones we do this, trend takes care of itself.

#### Checking the Differences needed

```{r}
ndiffs(cafe_win)
```

* We got ndiffs as 1 initially. 
* Let us first performs the differences for seasonal data and check ndiffs again.

#### Rechecking the differencing needed after performing the seasonal differencing

```{r}
ndiffs((diff(cafe_win,4))) #4 because it is quarterly
```

* As mentioned above, we can see that the ndiffs now becomes zero indicating that the trend component took care of itself.

#### Acf and Pacf plots

```{r}
tsdisplay(diff(cafe_win,4))
```

* In the Acf plot we can see that the p value is 1 and the q values is 1.
* So, the model takes values of 0,1 for p and q.
* In the Pacf plot, we can see that the P value is 1 and the Q value is 1.
* So, the model takes values of 0,1 for P and Q.
* However, the model also tests values other than the above to check if there is a better model with lesser AIC.

#### Arima Model

```{r}
arima_fit <- auto.arima(cafe_win,trace=TRUE, stepwise = FALSE )
arima_fit
```

* The system runs the auto arima and gives out the best model for which the AIC value is the lowest.
* We can see that the model tested multiple p,q,P,Q values other than the ones we predicted earlier.
* Finally the best model has p = 1, q = 0, P = 0, Q = 0.

#### Arima Model attributes

```{r}
attributes(arima_fit)
```

#### Plotting forecast of the Arima model with 99.5% confidence

```{r}
plot(forecast(arima_fit,h=5,level=c(99.5)), main = 'Forecast of the ARIMA model', xlab = 'Year', ylab = 'Expenditure')
```

## Residual Analysis

#### Acf of Residuals

```{r}
Acf(arima_fit$residuals)
```

* The Acf plot of the residuals shows white noice.
* None of the data passes the confidence interval limits.
* Which means that the forecast is a good forecast.

#### Box-Ljung test

```{r}
Box.test(residuals(arima_fit), lag=20, type="Ljung")
```

* The p-value is greater than 0.05 meaning the null hypothesis is supported.
* The residuals are random.

#### Residuals plot

```{r}
plot.ts(residuals(arima_fit))
```

* The residuals appear to be random.
* The residuals seem to the mean around zero.
* This suggests that it is a goof forecasting model.

#### Histogram plot of residuals

```{r}
hist(arima_fit$residuals)
```

* The histogram shows that the residuals have a mean of zero.
* The distribution is not a perfect normal distribution.
* The residuals are skewed to some extent.

#### TSDiagram of the ARIMA model

```{r}
tsdiag(arima_fit)
```

* The tsdiag() shows the overall residual analysis plots.
* The standardized residuals plot looks good. 2 residuals are very close to touch the z-score of 2 but did not touch the line.
* There are no significant values in the Acf plot.
* The p-values in Ljung-Box test are also above level, meaning the residuals are random.
* Overall, this ARIMA model is a good fit.





