---
title: "Midterm_BF"
author: "aa2569@scarletmail.rutgers.edu"
date: "2022-10-25"
output: html_document
---

```{r}
library(fpp)
library(fpp2)
library(TTR)
library(ggplot2)
library(readr)
NJ_MedianListingPrice_AllHomes <- read_csv(file = 'NJ_MedianListingPrice_AllHomes.csv')
NJ_Home_Raw <- NJ_MedianListingPrice_AllHomes$Value
NJ_Home_TS <- ts(NJ_Home_Raw,frequency = 12, start = c(1996,4))
```

## Plot and Inference

##### Q: Show a time series plot. 

```{r}
plot(NJ_Home_TS, main = 'Median home prices for House Listing in New Jersey', xlab = 'Year', ylab = 'Median home prices')
```

##### Q: Please summaries your observations of the times series plot

* The plot shows that there is an increasing trend in the median home prices starting from 1996 till around 2006. 
* From 2006 till 2012, there has been a decreasing trend in the home prices.
* From 2012, there has been steady increasing trend till the year 2017.
* The data however doesn't appear to show any seasonal variation.
* If we were to forecast the data, we should be considering the window from 2012 to 2017.

## Central Tendency

##### Q: What are the min, max, mean, median, 1st and 3rd Quartile values of the times series? 

```{r}
summary(NJ_Home_TS)
```

* The summary function above gives the min, max, mean, median, 1st and 3rd Quartile values of the times series.

##### Q: Show the box plot.
```{r}
boxplot(NJ_Home_TS, main = 'Boxplot for the Median House Prices Time Series')
hist(NJ_Home_TS)
```

##### Q: Can you summarize your observation about the time series from the summary stats and box plot? 

* The boxplot shows that there are no outliers in the data.
* The Median is more towards the third quartile.
* From summary, we can also see that the median value is greater than the mean for the time series.
* This means that the data is left skewed. This can be justified seeing the histogram above as well.

## Decomposition

##### Plot the decomposition of the time series.

```{r}
stl_decomp <- stl(NJ_Home_TS,s.window ="periodic")
plot(stl_decomp, main = 'Decomposition plot')
```

##### Is the times series seasonal?
* No, the time series do not show any seasonal variation.

##### Is the decomposition additive or multiplicative? 
* The decomposition is multiplicative. 
* Because, with time the trend seems to be also increasing in the data set.

##### If seasonal, what are the values of the seasonal monthly indices? 
##### For which month is the value of time series high and for which month is it low? 
##### Can you think of the reason behind the value being high in those months and low in those months?
* The data set doesn't have a seasonal variation. So the above three questions do not make sense for this time series. 
* We can check this hypothesis in the coming question.

##### Show the plot for time series adjusted for seasonality. Overlay this with the line for actual time series? Does seasonality have big fluctuations to the value of time series? 
```{r}
plot(NJ_Home_TS)
lines(seasadj(stl_decomp), col="Red")
```

* As expected, the seasonal adjusted plot exactly overlapped with the actual data.
* This states that our hypothesis is true and there appears to not have any seasonality in this time series

## Naïve Method

##### Output
```{r}
naive_forecast = naive(NJ_Home_TS)
plot(naive_forecast)
```

##### Perform Residual Analysis for this technique. 
###### Do a plot of residuals. What does the plot indicate?
```{r}
plot(naive_forecast$residuals)
```

* The residuals appear to have increasing positive values and then peaked in the year 2006 and then dipped and went on to be negative.
* The residuals had a negative peak in the year 2008 and then slowly evened out from the year 2013.
* We can test this hypothesis in the coming tests.

###### Do a Histogram plot of residuals. What does the plot indicate?
```{r}
hist(naive_forecast$residuals)
```
* The histogram appears to be normally distributed.
* But the values do not have a mean zero. The histogram appears to be skewed on one side.
* This means that the data is biased as the mean is not zero.

###### Do a plot of fitted values vs. residuals. What does the plot indicate? 

###### Do a plot of actual values vs. residuals. What does the plot indicate?

###### Do an ACF plot of the residuals? What does this plot indicate?

```{r}
Acf(naive_forecast$residuals)
```

* All the values of the Acf lot have crossed the confidence interval.
* Meaning that naive forecast is missing some main variables which we have missed our consideration for the forecast.

##### Print the 5 measures of accuracy for this forecasting technique

```{r}
accuracy(naive_forecast)
```

##### Forecast 
###### Time series value for next year. Show table and plot

```{r}
forecast(naive_forecast)
plot(forecast(naive_forecast))
```

##### Summarize this forecasting technique
###### How good is the accuracy?
* The ME, RMSE values are very high indicating that this method may not be the right one to go with.
* We can consider more forecasting techniques and check if the error values are less than this one.

###### What does it predict the value of time series will be in one year?
* Based on naive forecast, the predicted value of the time series in one year is 311600.

###### Other observation
* From 2013 to 2017 there is observed to be an increasing trend in the data. So, naive forecast may not be a right way to forecast.
* Rather, we can try naive method with drift component and that may yield us better forecast.


## Simple Moving Averages

##### Plot the graph for time series. 
##### Show the Simple Moving average of order 3 on the plot above in Red
##### Show the Simple Moving average of order 6 on the plot above in Blue
##### Show the Simple Moving average of order 9 on the plot above in Green

```{r}
mavg3_forecast = ma(NJ_Home_TS,order=3)
mavg6_forecast = ma(NJ_Home_TS,order=6)
mavg9_forecast = ma(NJ_Home_TS,order=9) 
plot(NJ_Home_TS, main = "Plot along with moving averages")
lines(mavg3_forecast, col="Red")
lines(mavg6_forecast, col="Blue")
lines(mavg9_forecast, col="Green")
```

##### (Bonus) show the forecast of next 12 months using one of the simple average order that you feel works best for time series
```{r}
plot(forecast(mavg9_forecast), main = "Forecasting for next months using order 9")
```

##### What are your observations of the plot as the moving average order goes up? 
* From the plots, it is observed that the higher the order we consider, the smoother the moviing average curve in the plot.
* It can be seen that the Green line above is the smoothest compared to Blue or Red lines.

## Simple Smoothing
##### Perform a simple smoothing forecast for next 12 months for the time series. 
###### What is the value of alpha?  What does that value signify? 
###### What is the value of initial state? 
###### What is the value of sigma?  What does the sigma signify?

```{r}
ses_data <- ses(NJ_Home_TS)
plot(ses_data)
attributes(ses_data)
```

##### Perform Residual Analysis for this technique. 
###### Do a plot of residuals. What does the plot indicate?

```{r}
plot(ses_data$residuals)
```

* The residuals appear to have increasing positive values and then peaked in the year 2005 and then dipped and went on to be negative.
* The residuals had a negative peak in the year between 2008 - 2009 and then slowly evened out from the year 2013.

###### Do a Histogram plot of residuals. What does the plot indicate?

```{r}
hist(ses_data$residuals)
```
* The histogram appears to be normally distributed.
* But the values do not have a mean zero. The histogram appears to be skewed on one side.
* If the residual histogram doesnot have the mean to be zero, it means the data is biased.

###### Do a plot of fitted values vs. residuals. What does the plot indicate? 
###### Do a plot of actual values vs. residuals. What does the plot indicate?

###### Do an ACF plot of the residuals? What does this plot indicate?

```{r}
Acf(ses_data$residuals)
```

* All the values of the Acf lot have crossed the confidence interval.
* Meaning that simple smoothing forecast is missing some main variables which we have missed our consideration for the forecast.

##### Print the 5 measures of accuracy for this forecasting technique

```{r}
accuracy(ses_data)
```

##### Forecast 
###### Time series value for next year. Show table and plot

```{r}
forecast(ses_data)
plot(forecast(ses_data))
```

##### Summarize this forecasting technique
###### How good is the accuracy?
* The ME, RMSE values are very high indicating that this method may not be the right one to go with.
* We can consider more forecasting techniques and check if the error values are less than this one.

###### What does it predict the value of time series will be in one year?
* Based on simple smoothing forecast, the predicted value of the time series in one year is 311599.9.

###### Other observation
* From 2013 to 2017 there is observed to be an increasing trend in the data. So, this forecast may not be a right way to forecast.
* We need to try other forecasting methods and check the accuracies in each case.

## Holt-Winters 
##### Perform Holt-Winters forecast for next 12 months for the time series.

```{r}
HW_forecast = HoltWinters(NJ_Home_TS)
plot(forecast(HW_forecast))
hw <- forecast(HW_forecast)
```

###### What is the value of alpha?  What does that value signify? 

```{r}
HW_forecast$alpha
```

* Alpha specifies the coefficient for the level smoothing in Holtwinters.
* Values near 1.0 mean that the latest value has more weight.

###### What is the value of beta? What does that value signify?

```{r}
HW_forecast$beta
```

* Beta specifies the coefficient for the trend smoothing in Holtwinters. 
* Values 1.0 means that the latest value has highest weight

###### What is the value of gamma? What does that value signify?

```{r}
HW_forecast$gamma
```

* Gamma specifies the coefficient for the seasonal smoothing in Holtwinters. 
* Values 1.0 means that the latest value has highest weight.

###### What is the value of initial states for the level, trend and seasonality? What do these values signify? 
###### What is the value of sigma?  What does the sigma signify?

##### Perform Residual Analysis for this technique. 
###### Do a plot of residuals. What does the plot indicate?

```{r}
plot(hw$residuals)
```

* The residuals appear to be random and also the mean looks to be near zero. We can check this with histogram.
* We can observe a couple of up and downs in 2009 and 2017. But even they did not show and growing residual pattern.

###### Do a Histogram plot of residuals. What does the plot indicate?

```{r}
hist(hw$residuals)
```
* The histogram appears to be normally distributed.
* And the mean is near zero. Indicating the data is not biased.
* Overall, comparing the previous forecasts, this forecast appears to be the best till now. 

###### Do a plot of fitted values vs. residuals. What does the plot indicate? 
###### Do a plot of actual values vs. residuals. What does the plot indicate?

###### Do an ACF plot of the residuals? What does this plot indicate?

```{r}
Acf(hw$residuals)
```

* Quite a few values of the Acf lot have crossed the confidence interval.
* The Acf crossed the confidence interval at 1, 2, 3, 4, 12, 24 lags.
* Which states that there are still few variables which we have missed in our forecast and that our forecast can still be improved.

##### Print the 5 measures of accuracy for this forecasting technique

```{r}
accuracy(hw)
```

##### Forecast 
###### Time series value for next year. Show table and plot

```{r}
forecast(HW_forecast)
plot(forecast(HW_forecast))
```

##### Summarize this forecasting technique
###### How good is the accuracy?
* The ME, RMSE values are quite low compared to any of our previous forecasts.
* Holwinters is a better forecast compared to naive and simple smoothing.

###### What does it predict the value of time series will be in one year?
* Based on HoltWinters forecast, the predicted value of the time series in one year is 331051.8.

###### Other observation
* Holtwinters appears to be the best forecast considering all the previous forecast methods.
* However, this forecast can still be improved as we can see that the residuals have crossed the confidence interval in the Acf plot.

## Accuracy Summary
##### Show a table of all the forecast method above with their accuracy measures.

```{r}
accuracy(naive_forecast)
accuracy(ses_data)
accuracy(hw)
```

##### Show the best and worst forecast method for the accuracy measure of your choice. Why did you choose that accuracy measure?

* To start with, there is nothing like best or worst forecast.
* Considering the accuracy data above, HoltWinters forecast seems to fit the time series the best as it has the least error values.
* And naive forecast seems to be the worst as it has the largest ME and RMSE values.

## Conclusion
##### Summarize your analysis of time series value over the time-period. 

* Upon initial observation, the data looked to have only trend and did not appear to show any seasonal variation.
* We have checked the hypothesis and have observed that there is no seasonal variation in the time series.
* Based on the three forecasting methods naive, simple smoothing, and HoltWinters, we can see that HoltWinters forecast provides to be the better forecasting method in this case.
* This is because the forecast fits perfectly and also the error values are quiet low for HoltWinters forecast.

##### Based on your analysis and forecast above, do you think the value of the time series will increase, decrease or stay flat over the next year? How about next 2 years?

* Based on the analysis and forecast, the time series will increase over the next year and the next 2 years.



  
 
  
  
  